It's Never Too Late To Learn\\ \emph{Using Online Machine Learning in Distributed Systems}

### 2016-01-22 writeups/msr_workshop/paper.tex
new intro

@@

Building a modern distributed service that guarantees service-level objectives
is hard. Complexity is introduced by the deployment environment and exacerbated
as the service evolves with time to accommodate hardware/environment changes or
support new features.
For example, to build a simple caching layer that guarantees a certain response
latency and throughput, developers need to take into account the number of
machines available to run that service, the network connecting these machines,
the storage hierarchy within each machine, the expected workload that the
caching layer will be servicing, and interference from other applications on the
network and machines.  A change in any one of these parameters could render the
service wildly inefficient.

Traditionally, this complexity is addressed by using highly-skilled developers
to build custom systems that are optimized for specific configurations and
objective functions.  Is there a general, principled approach to dealing with
complexity in distributed systems?
In this position paper, we propose a research project to study the use of Online
Machine Learning (ML) in distributed systems.  Hand-tuning the design and
configuration of distributed systems does not scale well with increased
complexity of software design, hardware architecture, deployment environment,
and strict service-level objectives.
To highlight this, here are some examples of where we think ML can help:

- contextual multi-armed bandits. idea is to provide a context, get an action,
and only observe reward for that action. This matches the dist systems world well. It will generate new
policies using ML that we could not come up with by just thinking, and may not
even be able to describe. We argue in favor of opacity and rely instead on the guarantees
provided by the ML optimization.

We describe some of hte unkowns we face and where ML can come in

@@

To be clear, ML has been used in some distributed systems. However, it has not
been used in a general way to make online decisions about core distributed
systems implementation details. We advocate that developers should focus on
high-level objectives of their system, and use online ML to make to compute the
specifics (such as policies for request routing, replica placement, or failure
monitoring) at run-time based on observed metrics.

@@

>>>>

\newcommand{\ignore}[1]{}

@@

Building a modern distributed service that guarantees service-level objectives
is hard. Complexity is introduced by the deployment environment and exacerbated
as the service evolves with time to accommodate hardware/environment changes or
support new features.
For example, to build a simple caching layer that guarantees a certain hit rate,
response latency, and throughput, developers need to take into account the expected
workload hitting the cache, the number of machines available to run it, the network
connecting these machines, the storage hierarchy within each machine, and interference
from other applications running on the network and machines.  A change in any one
of these factors could render the service inefficient.

Traditionally, this complexity is addressed by using highly-skilled developers
to build custom systems that are optimized for specific configurations and
objective functions.  Is there a general, principled approach for dealing with
complexity in distributed systems? Hand-tuning the design and configuration of a system
does not scale well with increased complexity and changes in software design, hardware
architecture, deployment environment, and service-level objectives.

We propose to use online learning to cope with this complexity. Specifically, we
view a {\em decision} made by a distributed system in the framework of
contextual online learning with partial feedback. This is a natural fit: we typically
have plenty of context surrounding a decision (which we may not know how to use), the decision
is made online, and we only receive feedback for the decision that was made. Thus, we can
replace a hand-designed policy with a new policy that has been optimized by a contextual
learning system. By supplying rich contexts to this policy, we can capture the complexity of
our environment without having to understand it; by optimizing the policy continuously, we can
can adapt to changes in the environment or service over time.

Machine learning has certainly been used in the past to optimize system decisions. Most
of this use has been restricted to offline settings with full information: for example,
supervised learning is often used to train models on data that has been annotated with
correct labels. Some systems work in an online setting with full information: for example,
a cache that applies an experts learning framework to dynamically switch between existing
(hand-designed) cache eviction policies, by running all policies in parallel and measuring
their performance. Systems that work in an online setting with partial information tend to
be for news or ad recommendation, but examples from systems infrastructure exist: for example,
caching dynamic query content [], tuning parameters of multicore data structures [], allocating
resources to cluster computing tasks [], and allocating servers to web applications [].
In all of these systems, the context used to make a decision, and the
decision itself, are {\em local} to a machine. This is achieved by either devising a decentralized
algorithm in the first place, or giving up on global optimality. Thus we are not aware of any
system that makes ``global'' decisions based on truly distributed architecture or state.

And yet we use hand-designed policies to make these decisions all the time, such as for request
routing, replica placement, replica selection, cloud resource allocation, failure recovery tasks, 
and others. What is it about distributed systems that makes it difficult to apply online learning
at standard decision points? This proposal identifies fundamental challenges posed by distributed
systems that must be analyzed and addressed in order to make online learning a standard tool for
cloud optimization.  These challenges hold even if we assume that the online learning system is
a black-box that delivers optimal decisions in real-time--in other words, they are fundamental
to the way we design distributed systems.  We begin by highlighting some of the unknown factors
that contribute to these challenges, and then list the challenges themselves.

@@

\ignore{
To be clear, ML has been used in some distributed systems. However, it has not
been used in a general way to make online decisions about core distributed
systems implementation details. We advocate that developers should focus on
high-level objectives of their system, and use online ML to make to compute the
specifics (such as policies for request routing, replica placement, or failure
monitoring) at run-time based on observed metrics.
}

@@

\ignore{
contextual multi-armed bandits. idea is to provide a context, get an action,
and only observe reward for that action. This matches the dist systems world well. It will generate new
policies using ML that we could not come up with by just thinking, and may not
even be able to describe. We argue in favor of opacity and rely instead on the guarantees
provided by the ML optimization.
}
